### 2015-05-13 whoop.tex
split optimisations and assumptions
updates

\medskip\noindent\textbf{Watchdog Race-Checking Instrumentation }
%
Data race detection is performed by introducing sets containing the locks that are consistently held by each shared variable and sets containing all shared variables that are read and written (see Section~\ref{sec:symbolicpairwise}). Such sets can be typically modeled in Boogie using maps. However, this requires the use of quantifiers to express invariants related to the contents of a set, such as emptiness. To avoid using quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog} race-checking instrumentation, which works as follows:

@@

\medskip\noindent\textbf{Watchdog Summarization }
%
Early on during the development of \whoop, we realized that scalability would be a serious issue. When we tried to apply \whoop on the r8169 RealTek ethernet driver (8000 lines of C code) with all functions fully inlined, we were running out of memory. One reason was that the C to Boogie translation produced hundreds of thousands of Boogie IVL code, which could not fit in memory. Another reason was that some entry points recursively call helper functions. Inherently, recursion does not work with inlining. To tackle these issues we developed \emph{watchdog summarization}, a novel summarization technique that allows us to analyze Linux drivers in a scalable fashion, but without destroying precision.

@@

The candidate invariants are based on the global variables that \whoop instruments to perform lockset analysis and race-checking (see Section~\ref{whoop:method}). The key idea behind watchdog summarization is that we automatically generate candidate invariants for each memory region in the granularity of watched accesses. This reduces the false positives from reporting errors if only e.g.\ a field in memory region \texttt{\$M.0} is racing, but the rest of the fields are not. An example of watchdog-based candidate post-conditions is the following:

@@

To generate the candidate invariants, \whoop performs an intra-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region. For example the previous \texttt{\$M.0} candidates would now be:

\begin{boogie}
procedure foo$ep1(dev: int);
  ensures _b$ls$entrypoint1$0
    ==> lock$0_in_LS_$M.0_$entrypoint1;
  ensures _b$ls$entrypoint1$1
    ==> lock$0_in_LS_$M.0_$entrypoint1;
  ensures _b$ac$entrypoint1$0
    ==> !WRITTEN_$M.0_$entrypoint1;
  ensures _b$ac$entrypoint1$1
    ==> !WRITTEN_$M.0_$entrypoint1;
  ...
\end{boogie}

@@

\medskip\noindent\textbf{Optimizations and Assumptions }
%
To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel interacts with its drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then exploit to make \whoop more precise.

@@

%We also optimize the use of function pointers. LLVM and SMACK do not perform context-sensitive translation, which means that a function that accepts a function pointer as an argument, can arbitrarily call any possible method that corresponds to that function pointer regardless of call site. To increase precision, Chauffeur analyses the AST of the original program and attempts to infer context-sensitive function pointer use information. This information is the used by \whoop to increase precision. To be sound, this optimization is conservative: if no information can be extracted for a function pointer at a call site, then \whoop uses the original over-approximated function.

Finally, we assume that the formal parameters of an entry point do not alias, and thus cannot race. This is a potentially unsound feature that can be turned off using a command line option. In our experience so far, though, we have not missed any data races because of this assumption. We also assume that our environmental model and domain-specific knowledge are correct and do not cause any bugs to be missed.

>>>>

\medskip\noindent\textbf{Watchdog Race-Checking Instrumentation }
%
Data race detection is performed by introducing sets containing the locks that are consistently held by each shared variable and sets containing all shared variables that are read and written (see Section~\ref{sec:symbolicpairwise}). Such sets can be typically modeled in Boogie using maps. However, this requires the use of quantifiers to express invariants related to the contents of a set, such as emptiness. To avoid using quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog race-checking instrumentation}, which works as follows:

@@

\medskip\noindent\textbf{Watchdog Summarization }
%
Early on during the development of \whoop, we realized that scalability would be a serious issue. When we tried to apply \whoop on the r8169 ethernet driver ($>$7000 lines of C code) with all functions fully inlined, we run out of memory. One reason was that the C to Boogie translation produced hundreds of thousands of Boogie IVL code, which could not fit in memory. Another reason was that some entry points recursively call helper functions. Inherently, recursion does not work well with inlining. To tackle these issues, we developed \emph{watchdog summarization}, a novel summarization technique that allows us to analyze drivers in a scalable fashion and without destroying precision.

@@

The candidate invariants are based on the aforementioned global watchdog variables that \whoop instruments for watchdog race-checking. The key idea behind watchdog summarization is to automatically generate candidate invariants for each memory region in the granularity of watched accesses. This reduces the false positives from reporting errors if only e.g.\ a field in memory region $\mathit{MR}$ is racing, but the rest of the fields in $\mathit{MR}$ are not racing.

To generate the candidate invariants, \whoop performs an inter-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region. For example the previous \texttt{\$M.0} candidates would now be:

@@

@@

\medskip\noindent\textbf{Optimizations }
%
To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel interacts with its drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then exploit to make \whoop more precise.

@@

\medskip\noindent\textbf{Assumptions }
%
We assume that the formal parameters of an entry point do not alias, and thus cannot race. This is a potentially unsound feature that can be turned off using a command line option. In our experience so far, though, we have not missed any data races because of this assumption. We also assume that our environmental model and domain-specific knowledge are correct and do not cause any bugs to be missed. Finally, we assume that the external tools we use as part of the \whoop toolchain are sound and bug-free.
